---
title: "Methods of Interpolation for Pattern Mining"
author: Caleb Koch
date: August 2016
output:
  revealjs::revealjs_presentation:
    incremental: true
    transition: slide
    theme: simple
    css: styleCSS.css
    self_contained: false
    reveal_plugins: ["notes"]
---

# Background - What is pattern mining? {.position}

##Pattern mining
- Similarities in sequential datasets
<br><br>

<div class="fragment">
### Pattern Mining Project
</div>

<div class="fragment">
Spatiotemporal moving object trajectories
</div>

<div class="fragment">
Large, noisy data
</div>

<div class="fragment">
Hidden, nontrivial patterns and relationships (including associations between trajectories)
</div>

##mSEQUITUR
```{r echo=FALSE, fig.height=5}
library(png)
library(grid)
img <- readPNG("FlowDiagram.png")
 grid.raster(img)
```

##Data Collection
GeoLife Data

39.906631,116.385564,0,492,40097.5864583333,2009-10-11,14:04:30

```{r, echo=FALSE, warning=FALSE, fig.height=3.5, width=3}

v = read.table("20080927233805.plt", header = FALSE, quote = "\"", skip = 6, sep=",")
x1 <- v$V2
y1 <- v$V1

library(ggplot2)

ggplot(aes(x = x1, y = y1, color="black"), data=v) + geom_point(size=0.5, color="#070688") + labs(x = "Longitude", y = "Latitude") + ggtitle("Example Trajectory")
```

##Alternative View
```{r, echo = FALSE, fig.height=6, fig.width=9}
library(scatterplot3d)

v = read.table("20080927233805.plt", header = FALSE, quote = "\"", skip = 6, sep=",")
x1 <- v$V1
y1 <- v$V2
z1 <- v$V5

scatterplot3d(x1, y1, z1, xlab = "Latitude", ylab = "Longitude", zlab = "Time",  label.tick.marks=FALSE)
```

##Example of a Time Series
```{r, echo=FALSE, warning=FALSE, fig.height=6}
ts <- scan("~/FinalPresentation/20080927233805_TS.plt", what = integer(), sep = ",")
plot(ts, ylab = "Hilbert Value", pch=16, cex = 0.5)
```

## Problem Overview

<div class="fragment">
Identify potential interpolation methods
</div>

<div class="fragment">
Assess interpolation methods
</div>

<div class="fragment">
Determine effects at different preprocessing stages
</div>


## Motivation
- Linearly interpolating trajectories yields large errors
<div class="fragment">
```{r, fig=TRUE, echo=FALSE, cache=TRUE, fig.height=5, fig.align='center'}
linearData3 = read.table("C:/test/lin3.txt", header = FALSE, quote = "\"", skip = 0, sep="\t")
xl3 <- linearData3$V1
yl3 <- linearData3$V2

trajData3 = read.table("C:/test/traj3.txt")
xt3 <- trajData3$V1
yt3 <- trajData3$V2

library(ggplot2)

t4 <- ggplot(aes(x = xl3, y = yl3, color="black"), data=linearData3) + geom_point(size=0.5, color="#070688") + geom_point(size=0.5, aes(x = xt3, y = yt3), color="#898da9", data=trajData3)  + geom_line(aes(x = xt3, y = yt3), color="#898da9", data=trajData3) + labs(x = "Longitude", y = "Latitude") + ggtitle("Sample Trajectory")

t4
```
</div>

# Search and Experimentation - How do we compare interpolation methods? {.position}

## Initial Search
- Pre-existing methods of interpolation
    - Local versus global (problems associated with distinction)
- Development of velocity-based model
    - Exploits acceleration and directional change to interpolate new points
    - Computationally intensive
    - Parameter optimization

##Interpolation Methods to Test
- Nearest Neighbor
- Linear
- Cubic Spline
- Polynomial
- Local regression (LOESS)

##Standard Experimentation
- Standard framework to assess accuracy across models (at different preprocessing stages)
<br><br>
<div class="fragment">
**General outline**:
<ol>
  <li>Select a sample of trajectories and generate artificial gaps</li>
  <li>Use interpolation methods to fill gaps in both trajectories and time series</li>
  <li>Run error analysis</li>
  <li>Compare errors by calculating representative inaccuracies</li>
</ol>
</div>

## Artificial Gaps: Example (563 KB)

```{r, echo=FALSE, fig=TRUE, cache=TRUE, warning=FALSE}
orig <- read.csv("C:/test/trajectory2/20090120055702.plt", sep=",", skip = 6)
f <- read.csv("C:/test/trajectory2/20090120055702_50.plt", sep=",", skip = 6)
o <- read.csv("C:/test/trajectory2/20090120055702_1000.plt", sep=",", skip = 6)
t <- read.csv("C:/test/trajectory2/20090120055702_2000.plt", sep=",", skip = 6)

x1 <- orig$X116.322061
y1 <- orig$X40.010435

x2 <- f$X116.322061
y2 <- f$X40.010435

x3 <- o$X116.322061
y3 <- o$X40.010435

x4 <- t$X116.322061
y4 <- t$X40.010435

library(ggplot2)
# qplot(trajData$V1, trajData$V2, ylab = "Latitude", xlab = "Longitude")

t1 <- ggplot(aes(x = x1, y = y1, color="black"), data=orig) + geom_point(size=0.5, color="#070688") + labs(x = "Longitude", y = "Latitude") + ggtitle("Original Trajectory")

t2 <- ggplot(aes(x = x2, y = y2, color="black"), data=f) + geom_point(size=0.5, color="#070688") + labs(x = "Longitude", y = "Latitude") + ggtitle("First Gap Length (50)")

t3 <- ggplot(aes(x = x3, y = y3, color="black"), data=o) + geom_point(size=0.5, color="#070688") + labs(x = "Longitude", y = "Latitude") + ggtitle("Second Gap Length (1000)")

t4 <- ggplot(aes(x = x4, y = y4, color="black"), data=t) + geom_point(size=0.5, color="#070688") + labs(x = "Longitude", y = "Latitude") + ggtitle("Third Gap Length (2000)")

library(grid)
library(gridExtra)
grid.arrange(t1, t2, t3, t4, ncol=2)
```

# Results - What did our analysis show? {.position}

## Performance comparison

```{r, echo=FALSE, fig=TRUE, cache=TRUE, warning=FALSE, fig.width=11, fig.height=7}
CubicErrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/CubicErrorTraj.plt", header=FALSE)
colnames(CubicErrorTraj) <- c("file", "error", "errorSquared", "length")

LinerrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/LinerrorTraj.plt", header=FALSE)
colnames(LinerrorTraj) <- c("file", "error", "errorSquared", "length")

NNerrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/NNerrorTraj.plt", header=FALSE)
colnames(NNerrorTraj) <- c("file", "error", "errorSquared", "length")

CubicTSError <- read.csv("C:/test/trajectory2/Error_Analysis/CubicTSError.plt", header=FALSE)
colnames(CubicTSError) <- c("file", "error", "errorSquared", "length")

LinerrorTS <- read.csv("C:/test/trajectory2/Error_Analysis/LinerrorTS.plt", header=FALSE)
colnames(LinerrorTS) <- c("file", "error", "errorSquared", "length")

NNerrorTS <- read.csv("C:/test/trajectory2/Error_Analysis/NNerrorTS.plt", header=FALSE)
colnames(NNerrorTS) <- c("file", "error", "errorSquared", "length")

totalDistance <- read.csv("C:/test/trajectory2/Error_Analysis/totalDistance.plt", header=FALSE)
colnames(totalDistance) <- c("trajectory", "timeSeries", "file")

LoessErrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/LoessErrorTraj.plt", header = FALSE)
colnames(LoessErrorTraj) <- c("file", "error", "errorSquared", "length")


LinerrorTS <- lapply(LinerrorTS, function(x){
  gsub("_TS", "", x)
})

NNerrorTS <- lapply(NNerrorTS, function(x){
  gsub("_TS", "", x)
})

LinerrorTraj <- lapply(LinerrorTraj, function(x){
  gsub("_LinInt_Traj.plt", "", x)
})

# Format CubicErrorTraj

CubicErrorTraj <- lapply(CubicErrorTraj, function(x){
  sub("_", "@", x)
})

test <- data.frame(do.call('rbind', strsplit(as.character(CubicErrorTraj$file),'_',fixed = TRUE)))

test <- lapply(test, function(x){
  sub("@", "_", x)
})

CubicErrorTraj$file <- test$X1


# Format LoessErrorTraj -----------------------------------------------------------

LoessErrorTraj <- lapply(LoessErrorTraj, function(x){
  sub("_", "@", x)
})

test <- data.frame(do.call('rbind', strsplit(as.character(LoessErrorTraj$file),'_',fixed = TRUE)))

test <- lapply(test, function(x){
  sub("@", "_", x)
})

LoessErrorTraj$file <- test$X1

# Format CubicTSError

CubicTSError <- lapply(CubicTSError, function(x){
  sub("_", "@", x)
})

CubicTSError <- lapply(CubicTSError, function(x){
  sub("_", "@", x)
})

test <- data.frame(do.call('rbind', strsplit(as.character(CubicTSError$file),'_',fixed = TRUE)))

test <- lapply(test, function(x){
  sub("@TS@", "_", x)
})

CubicTSError$file <- test$X1

# Merge Files
NNerrorTS <- merge(totalDistance, NNerrorTS, by.x = "file", by.y = "file")
NNerrorTraj <- merge(totalDistance, NNerrorTraj, by.x = "file", by.y = "file")
CubicTSError <- merge(totalDistance, CubicTSError, by.x = "file", by.y = "file")
CubicErrorTraj <- merge(totalDistance, CubicErrorTraj, by.x = "file", by.y = "file")
LinerrorTraj <- merge(totalDistance, LinerrorTraj, by.x = "file", by.y = "file")
LinerrorTS <- merge(totalDistance, LinerrorTS, by.x = "file", by.y = "file")
LoessErrorTraj <- merge(totalDistance, LoessErrorTraj, by.x = "file", by.y = "file")

# Add Gap Lengths

test <- data.frame(do.call('rbind', strsplit(as.character(NNerrorTS$file),'_',fixed = TRUE)))
NNerrorTS$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(NNerrorTraj$file),'_',fixed = TRUE)))
NNerrorTraj$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(CubicTSError$file),'_',fixed = TRUE)))
CubicTSError$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(CubicErrorTraj$file),'_',fixed = TRUE)))
CubicErrorTraj$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(LinerrorTraj$file),'_',fixed = TRUE)))
LinerrorTraj$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(LinerrorTS$file),'_',fixed = TRUE)))
LinerrorTS$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(LoessErrorTraj$file),'_',fixed = TRUE)))
LoessErrorTraj$Gaps <- test$X2

# Add inaccuracy
NNerrorTS$error <- as.numeric(as.character(NNerrorTS$error))
NNerrorTraj$error <- as.numeric(as.character(NNerrorTraj$error))
CubicTSError$error <- as.numeric(as.character(CubicTSError$error))
CubicErrorTraj$error <- as.numeric(as.character(CubicErrorTraj$error))
LinerrorTraj$error <- as.numeric(as.character(LinerrorTraj$error))
LinerrorTS$error <- as.numeric(as.character(LinerrorTS$error))
LoessErrorTraj$error <- as.numeric(as.character(LoessErrorTraj$error))

NNerrorTS$timeSeries <- as.numeric(as.character(NNerrorTS$timeSeries))
CubicTSError$timeSeries <- as.numeric(as.character(CubicTSError$timeSeries))
LinerrorTS$timeSeries <- as.numeric(as.character(LinerrorTS$timeSeries))

NNerrorTraj$trajectory <- as.numeric(as.character(NNerrorTraj$trajectory))
CubicErrorTraj$trajectory <- as.numeric(as.character(CubicErrorTraj$trajectory))
LinerrorTraj$trajectory <- as.numeric(as.character(LinerrorTraj$trajectory))
LoessErrorTraj$trajectory <- as.numeric(as.character(LoessErrorTraj$trajectory))

NNerrorTS$Inaccuracy <- NNerrorTS$error / NNerrorTS$timeSeries
CubicTSError$Inaccuracy <- CubicTSError$error / CubicTSError$timeSeries
LinerrorTS$Inaccuracy <- LinerrorTS$error / LinerrorTS$timeSeries

NNerrorTraj$Inaccuracy <- NNerrorTraj$error / NNerrorTraj$trajectory
CubicErrorTraj$Inaccuracy <- CubicErrorTraj$error / CubicErrorTraj$trajectory
LinerrorTraj$Inaccuracy <- LinerrorTraj$error / LinerrorTraj$trajectory
LoessErrorTraj$Inaccuracy <- LoessErrorTraj$error / LoessErrorTraj$trajectory


# Convert Gaps to numeric
NNerrorTS$Gaps <- as.numeric(as.character(NNerrorTS$Gaps))
NNerrorTraj$Gaps <- as.numeric(as.character(NNerrorTraj$Gaps))
CubicTSError$Gaps <- as.numeric(as.character(CubicTSError$Gaps))
CubicErrorTraj$Gaps <- as.numeric(as.character(CubicErrorTraj$Gaps))
LinerrorTraj$Gaps <- as.numeric(as.character(LinerrorTraj$Gaps))
LinerrorTS$Gaps <- as.numeric(as.character(LinerrorTS$Gaps))
LoessErrorTraj$Gaps <- as.numeric(as.character(LoessErrorTraj$Gaps))

# Get average of inaccuracies
x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(NNerrorTS, NNerrorTS$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
NNerrorTS2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(NNerrorTraj, NNerrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
NNerrorTraj2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(CubicTSError, CubicTSError$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
CubicTSError2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(CubicErrorTraj, CubicErrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
CubicErrorTraj2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(LinerrorTraj, LinerrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
LinerrorTraj2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(LinerrorTS, LinerrorTS$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
LinerrorTS2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(LoessErrorTraj, LoessErrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
LoessErrorTraj2 <- df

#Visualize Total Error
library(ggplot2)

df <- data.frame(NNerrorTS2$x, NNerrorTS2$y, NNerrorTraj2$y, CubicErrorTraj2$y, CubicTSError2$y, LinerrorTS2$y, LinerrorTraj2$y, LoessErrorTraj2$y)


ggplot(df, aes(NNerrorTS2$x, y=value, color=Method)) + labs(x="Gap Length", y="Inaccuracy")  + geom_point(aes(y=LinerrorTraj2$y, col="Linear (Trajectory)"), size=1.5) +geom_point(aes(y=CubicErrorTraj2$y,col="Cubic (Trajectory)"), size=1.5)+ geom_point(aes(y=NNerrorTraj2$y,col="N. Neighbor (Trajectory)"), size=1.5) + geom_point(aes(y=LinerrorTS2$y, col="Linear (TS)"), size=1.5) + geom_point(aes(y=NNerrorTS2$y,col="N. Neighbor (TS)"), size=1.5) + geom_point(aes(y=CubicTSError2$y,col="Cubic (TS)"), size=1.5) + geom_point(aes(y=LoessErrorTraj2$y,col="LOESS (Trajectory)"), size=1.5) + geom_line(aes(x = LinerrorTraj2$x, y = LinerrorTraj2$y), color="#53B400", data=LinerrorTraj2, size=0.5) + geom_line(aes(x = NNerrorTraj2$x, y = NNerrorTraj2$y), color="#A58AFF", data=NNerrorTraj2, size=0.5) + geom_line(aes(x = LinerrorTS2$x, y = LinerrorTS2$y), color="#00C094", data=LinerrorTS2, size=0.5) + geom_line(aes(x = NNerrorTS2$x, y = NNerrorTS2$y), color="#FB61D7", data=NNerrorTS2, size=0.5) + geom_line(aes(x = CubicTSError2$x, y = CubicTSError2$y), color="#C49A00", data = CubicTSError2, size = 0.5) + geom_line(aes(x = CubicErrorTraj2$x, y = CubicErrorTraj2$y), color="#F8766D", data = CubicErrorTraj2, size = 0.5) + geom_line(aes(x = LoessErrorTraj2$x, y = LoessErrorTraj2$y), color="#00B6EB", data = LoessErrorTraj2, size = 0.5)
```

## Comparison (excluding outlier results)

```{r, echo=FALSE, fig=TRUE, cache=TRUE, warning=FALSE, fig.width=11, fig.height=7}
CubicErrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/CubicErrorTraj.plt", header=FALSE)
colnames(CubicErrorTraj) <- c("file", "error", "errorSquared", "length")

LinerrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/LinerrorTraj.plt", header=FALSE)
colnames(LinerrorTraj) <- c("file", "error", "errorSquared", "length")

NNerrorTraj <- read.csv("C:/test/trajectory2/Error_Analysis/NNerrorTraj.plt", header=FALSE)
colnames(NNerrorTraj) <- c("file", "error", "errorSquared", "length")

CubicTSError <- read.csv("C:/test/trajectory2/Error_Analysis/CubicTSError.plt", header=FALSE)
colnames(CubicTSError) <- c("file", "error", "errorSquared", "length")

LinerrorTS <- read.csv("C:/test/trajectory2/Error_Analysis/LinerrorTS.plt", header=FALSE)
colnames(LinerrorTS) <- c("file", "error", "errorSquared", "length")

NNerrorTS <- read.csv("C:/test/trajectory2/Error_Analysis/NNerrorTS.plt", header=FALSE)
colnames(NNerrorTS) <- c("file", "error", "errorSquared", "length")

totalDistance <- read.csv("C:/test/trajectory2/Error_Analysis/totalDistance.plt", header=FALSE)
colnames(totalDistance) <- c("trajectory", "timeSeries", "file")


LinerrorTS <- lapply(LinerrorTS, function(x){
  gsub("_TS", "", x)
})

NNerrorTS <- lapply(NNerrorTS, function(x){
  gsub("_TS", "", x)
})

LinerrorTraj <- lapply(LinerrorTraj, function(x){
  gsub("_LinInt_Traj.plt", "", x)
})

# Format CubicErrorTraj

CubicErrorTraj <- lapply(CubicErrorTraj, function(x){
  sub("_", "@", x)
})

test <- data.frame(do.call('rbind', strsplit(as.character(CubicErrorTraj$file),'_',fixed = TRUE)))

test <- lapply(test, function(x){
  sub("@", "_", x)
})

CubicErrorTraj$file <- test$X1

# Format CubicTSError

CubicTSError <- lapply(CubicTSError, function(x){
  sub("_", "@", x)
})

CubicTSError <- lapply(CubicTSError, function(x){
  sub("_", "@", x)
})

test <- data.frame(do.call('rbind', strsplit(as.character(CubicTSError$file),'_',fixed = TRUE)))

test <- lapply(test, function(x){
  sub("@TS@", "_", x)
})

CubicTSError$file <- test$X1

# Merge Files
NNerrorTS <- merge(totalDistance, NNerrorTS, by.x = "file", by.y = "file")

NNerrorTraj <- merge(totalDistance, NNerrorTraj, by.x = "file", by.y = "file")

CubicTSError <- merge(totalDistance, CubicTSError, by.x = "file", by.y = "file")

CubicErrorTraj <- merge(totalDistance, CubicErrorTraj, by.x = "file", by.y = "file")

LinerrorTraj <- merge(totalDistance, LinerrorTraj, by.x = "file", by.y = "file")

LinerrorTS <- merge(totalDistance, LinerrorTS, by.x = "file", by.y = "file")


# Add Gap Lengths

test <- data.frame(do.call('rbind', strsplit(as.character(NNerrorTS$file),'_',fixed = TRUE)))
NNerrorTS$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(NNerrorTraj$file),'_',fixed = TRUE)))
NNerrorTraj$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(CubicTSError$file),'_',fixed = TRUE)))
CubicTSError$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(CubicErrorTraj$file),'_',fixed = TRUE)))
CubicErrorTraj$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(LinerrorTraj$file),'_',fixed = TRUE)))
LinerrorTraj$Gaps <- test$X2

test <- data.frame(do.call('rbind', strsplit(as.character(LinerrorTS$file),'_',fixed = TRUE)))
LinerrorTS$Gaps <- test$X2

# Add inaccuracy
NNerrorTS$error <- as.numeric(as.character(NNerrorTS$error))
NNerrorTraj$error <- as.numeric(as.character(NNerrorTraj$error))
CubicTSError$error <- as.numeric(as.character(CubicTSError$error))
CubicErrorTraj$error <- as.numeric(as.character(CubicErrorTraj$error))
LinerrorTraj$error <- as.numeric(as.character(LinerrorTraj$error))
LinerrorTS$error <- as.numeric(as.character(LinerrorTS$error))

NNerrorTS$timeSeries <- as.numeric(as.character(NNerrorTS$timeSeries))
CubicTSError$timeSeries <- as.numeric(as.character(CubicTSError$timeSeries))
LinerrorTS$timeSeries <- as.numeric(as.character(LinerrorTS$timeSeries))

NNerrorTraj$trajectory <- as.numeric(as.character(NNerrorTraj$trajectory))
CubicErrorTraj$trajectory <- as.numeric(as.character(CubicErrorTraj$trajectory))
LinerrorTraj$trajectory <- as.numeric(as.character(LinerrorTraj$trajectory))

NNerrorTS$Inaccuracy <- NNerrorTS$error / NNerrorTS$timeSeries
CubicTSError$Inaccuracy <- CubicTSError$error / CubicTSError$timeSeries
LinerrorTS$Inaccuracy <- LinerrorTS$error / LinerrorTS$timeSeries

NNerrorTraj$Inaccuracy <- NNerrorTraj$error / NNerrorTraj$trajectory
CubicErrorTraj$Inaccuracy <- CubicErrorTraj$error / CubicErrorTraj$trajectory
LinerrorTraj$Inaccuracy <- LinerrorTraj$error / LinerrorTraj$trajectory


# Convert Gaps to numeric
NNerrorTS$Gaps <- as.numeric(as.character(NNerrorTS$Gaps))
NNerrorTraj$Gaps <- as.numeric(as.character(NNerrorTraj$Gaps))
CubicTSError$Gaps <- as.numeric(as.character(CubicTSError$Gaps))
CubicErrorTraj$Gaps <- as.numeric(as.character(CubicErrorTraj$Gaps))
LinerrorTraj$Gaps <- as.numeric(as.character(LinerrorTraj$Gaps))
LinerrorTS$Gaps <- as.numeric(as.character(LinerrorTS$Gaps))

# Get average of inaccuracies
x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(NNerrorTS, NNerrorTS$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
NNerrorTS2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(NNerrorTraj, NNerrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
NNerrorTraj2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(CubicTSError, CubicTSError$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
CubicTSError2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(CubicErrorTraj, CubicErrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
CubicErrorTraj2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(LinerrorTraj, LinerrorTraj$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
LinerrorTraj2 <- df

x <- c(seq(from = 50, to = 2000, by = 50))
y <- x+2000
df <- data.frame(x, y)
for(i in seq(from = 50, to = 2000, by = 50)){
  test2 <- subset(LinerrorTS, LinerrorTS$Gaps==i)
  tempVal <- i+2000
  df[df==tempVal] <- mean(test2$Inaccuracy)
}
LinerrorTS2 <- df


#Visualize Total Error
library(ggplot2)

df <- data.frame(NNerrorTS2$x, NNerrorTS2$y, NNerrorTraj2$y, CubicErrorTraj2$y, CubicTSError2$y, LinerrorTS2$y, LinerrorTraj2$y)

ggplot(df, aes(NNerrorTS2$x, y=value, color=Method)) + labs(x="Gap Length", y="Inaccuracy") + geom_point(aes(y=LinerrorTraj2$y, col="Linear (Trajectory)"), size=1.5) + geom_point(aes(y=NNerrorTraj2$y,col="N. Neighbor (Trajectory)"), size=1.5) + geom_point(aes(y=LinerrorTS2$y, col="Linear (TS)"), size=1.5) + geom_point(aes(y=NNerrorTS2$y,col="N. Neighbor (TS)"), size=1.5) + geom_point(aes(y=CubicTSError2$y,col="Cubic (TS)"), size=1.5) + geom_line(aes(x = LinerrorTraj2$x, y = LinerrorTraj2$y), color="#C49A00", data=LinerrorTraj2, size=0.5) + geom_line(aes(x = NNerrorTraj2$x, y = NNerrorTraj2$y), color="#00B6EB", data=NNerrorTraj2, size=0.5) + geom_line(aes(x = LinerrorTS2$x, y = LinerrorTS2$y), color="#00C094", data=LinerrorTS2, size=0.5) + geom_line(aes(x = NNerrorTS2$x, y = NNerrorTS2$y), color="#FB61D7", data=NNerrorTS2, size=0.5) + geom_line(aes(x = CubicTSError2$x, y = CubicTSError2$y), color="#F8766D", data = CubicTSError2, size = 0.5)
```


<aside class="notes">
- Each method performed better when applied to time series data
- Error increases with gap length (exception: cubic trajectory)
- Linear and nearest neighbor returned similar results
- Nearest neighbor interpolation with time series data returned lowest errors
</aside>

# Concluding Remarks - What did we learn? {.position}


## Significance of work

- Standardized experimental methodology
    - Reproducibility and comparative analysis
- Implementation of procedures (accessibility of code)
- Documentation
    - Adaptability and mutability
- Data
    - Analysis
    - Availability
- Relation to pattern mining project
    - Comparative basis for performance analysis
- Use of nearest neighbor interpolation
- Development of global interpolation method

## Considerations for future research

- Automatic parameter estimation and model refinement
- Linearization via Azimuths (alternate preprocessing steps)
- Adaptive interpolation
- Analyzing other components of data

##Experience {.columns-2}
<div style="float: left; width: 50%;">
- R
    - Data analysis and visualization
    - RStudio
    - R Markdown
- Java
    - EGit
    - Dependency management
    - Software development
</div>

<div style="float: right; width: 50%;">
- Beamer
    - Beamer and R with Sweave
    - Typesetting pseudocode
- Rendering github flavored markdown docs with pandoc
- Presentations with R Markdown and reveal.js (rendered with knitr)
- Interpolation
</div>

## {data-background="cornell_6.jpg"}

## {data-background="winter2.jpg"}

## Acknowledgements {.position}

<aside class="notes">
  Academy of Applied Science, the Army Educational Outreach Program, the Geospatial Research Laboratory, my mentor and all of those at GRL who helped make this internship possible
</aside>
---




<!-- -->
<!-- <div class="center"> -->
<!-- ```{r, echo=FALSE} -->
<!-- library(diagram) -->
<!-- library(shape) -->
<!-- par(mar = c(1, 1, 1, 1), mfrow = c(1, 1, 1)) -->
<!-- names <- c("A", "B", "C") -->
<!-- M <- matrix(nrow = 3, ncol = 1, byrow = FALSE, data = 0) -->
<!-- M[3, 2] <- M[2, 1] <-"" -->
<!-- plotmat(M, pos = c(1, 1, 1), name = names, box.type = "square") -->
<!-- ``` -->
<!-- </div> -->

<!-- ## -->
<!-- ```{r, echo=FALSE, figure=FALSE} -->
<!-- library(DiagrammeR) -->
<!-- the_nodes <-  -->
<!--   create_nodes( -->
<!--     nodes = 1:5, -->
<!--     label = c("Data Collection","Linearization","Discretization","Grammar Induction","Motif Discovery")) -->
<!-- the_edges <-  -->
<!--   create_edges( -->
<!--     from = c(1, 2, 3, 4), -->
<!--     to   = c(2, 3, 4, 5)) -->
<!-- the_graph <-  -->
<!--   create_graph( -->
<!--     nodes_df = the_nodes, -->
<!--     edges_df = the_edges) -->

<!-- render_graph(create_graph(edges_df = the_edges,nodes_df = the_nodes),width = 800, height = 700) -->
<!-- ``` -->
<!-- --> 


<!-- ggplot(df, aes(NNerrorTS2$x, y=value, color=Method)) + labs(x="Gap Length", y="Inaccuracy")  + geom_point(aes(y=LinerrorTraj2$y, col="Linear (Trajectory)"), size=1.5) +geom_point(aes(y=CubicErrorTraj2$y,col="Cubic (Trajectory)"), size=1.5)+ geom_point(aes(y=NNerrorTraj2$y,col="N. Neighbor (Trajectory)"), size=1.5) + geom_point(aes(y=LinerrorTS2$y, col="Linear (TS)"), size=1.5) + geom_point(aes(y=NNerrorTS2$y,col="N. Neighbor (TS)"), size=1.5) + geom_point(aes(y=CubicTSError2$y,col="Cubic (TS)"), size=1.5) + geom_point(aes(y=LoessErrorTraj2$y,col="LOESS (Trajectory)"), size=1.5) + geom_line(aes(x = LinerrorTraj2$x, y = LinerrorTraj2$y), color="#00BA38", data=LinerrorTraj2, size=0.5) + geom_line(aes(x = NNerrorTraj2$x, y = NNerrorTraj2$y), color="#619CFF", data=NNerrorTraj2, size=0.5) + geom_line(aes(x = LinerrorTS2$x, y = LinerrorTS2$y), color="#00BFC4", data=LinerrorTS2, size=0.5) + geom_line(aes(x = NNerrorTS2$x, y = NNerrorTS2$y), color="#FB61D7", data=NNerrorTS2, size=0.5) + geom_line(aes(x = CubicTSError2$x, y = CubicTSError2$y), color="#C49A00", data = CubicTSError2, size = 0.5) + geom_line(aes(x = CubicErrorTraj2$x, y = CubicErrorTraj2$y), color="#F8766D", data = CubicErrorTraj2, size = 0.5) + geom_line(aes(x = LoessErrorTraj2$x, y = LoessErrorTraj2$y), color="#00BFC4", data = LoessErrorTraj2, size = 0.5) + scale_fill_manual(breaks=c("Cubic (Trajectory)","LOESS (Trajectory)","N. Neighbor (TS)", "Linear (Trajectory)", "Cubic (TS)", "N. Neighbor (TS)", "Linear (TS)"), labels=c("Cubic (Trajectory)","LOESS (Trajectory)","N. Neighbor (TS)", "Linear (Trajectory)", "Cubic (TS)", "N. Neighbor (TS)", "Linear (TS)"))-->

